Autopolyploid establishment through polygenic adaptation

### 2025-04-22 doc/paper/paper.tex
update

    \begin{document}
    \title{Autopolyploid establishment through polygenic adaptation}
    %\author[1]{Arthur Zwaenepoel\thanks{arthur.zwaenepoel@uantwerpen.be}}
    %\affil[1]{\footnotesize Department of Biology, University of Antwerp, 2020 Antwerp, Belgium}
    \author{}
    \date{\vspace{-5ex}}
    \maketitle
    \begin{abstract}
    We define the infinitesimal model of quantitative genetics for the inheritance
        of an additive quantitative trait in a mixed-ploidy population consisting
        of diploid, triploid and autotetraploid individuals producing haploid and
        diploid gametes.
        We implement efficient simulation methods and use these to study the
        quantitative genetics of mixed-ploidy populations and the establishment of
        autotetraploids in a new habitat.
        We show that, when migration from the source population is rare,
        autotetraploids are more likely to establish in the new habitat 
        than diploids under a very broad range of conditions, but that this
        is unlikely to sufficiently counter the scarcity of tetraploid founders
        when the source is predominantly diploid.
        We assess in more detail how minority cytotype exclusion interacts with
        migration load in the establishment process and evaluate the impact of
        additional sources of prezygotic isolation, specifically selfing and
        assortative mating, on the relative establishment probabilities of the
        different cytotypes in the presence of maladaptive migration.
        In the discussion we consider how inbreeding depression may impact our
        findings.

@@

    Individual-based simulations for the mixed-ploidy infinitesimal model were
    implemented in Julia \citep{julia}.
    Documented code and simulation notebooks are available at \url{[anonymized]}.
%\url{https://github.com/arzwa/InfGenetics}.

@@

%\section*{Acknowledgements}
%
%I thank Christelle Fraïsse for encouragement and feedback. Thanks to Quinten
%Bafort, Felipe Kauai and Frederik Mortier for comments on an early draft
%version.
%I acknowledge funding from the Research Foundation -- Flanders (FWO, Junior
%Postdoctoral Fellowship 1272625N).

>>>>

    \begin{document}
    \title{Autopolyploid establishment through polygenic adaptation}
    \author[1]{Arthur Zwaenepoel\thanks{arthur.zwaenepoel@uantwerpen.be}}
    \affil[1]{\footnotesize Department of Biology, University of Antwerp, 2020 Antwerp, Belgium}
    %\author{}
    \date{\vspace{-5ex}}
    \maketitle
    \begin{abstract}
    We define the infinitesimal model of quantitative genetics for the inheritance
        of an additive quantitative trait in a mixed-ploidy population consisting
        of diploid, triploid and autotetraploid individuals producing haploid and
        diploid gametes.
        We implement efficient simulation methods and use these to study the
        quantitative genetics of mixed-ploidy populations and the establishment of
        autotetraploids in a new habitat.
        We show that, when migration from the source population is rare,
        autotetraploids are more likely to establish in the new habitat 
        than diploids under a very broad range of conditions, but that this
        is unlikely to sufficiently counter the scarcity of tetraploid founders
        when the source is predominantly diploid.
        We assess in more detail how minority cytotype exclusion interacts with
        migration load in the establishment process and evaluate the impact of
        additional sources of prezygotic isolation, specifically selfing and
        assortative mating, on the relative establishment probabilities of the
        different cytotypes in the presence of maladaptive migration.
        In the discussion we consider how inbreeding depression may impact our
        findings.

@@

    Individual-based simulations for the mixed-ploidy infinitesimal model were
    implemented in Julia \citep{julia}.
    %Documented code and simulation notebooks are available at \url{[anonymized]}.
    Documented code and simulation notebooks are available at
    \url{https://github.com/arzwa/InfGenetics}.

@@

\section*{Funding}

I acknowledge funding from the Research Foundation -- Flanders (FWO, Junior
Postdoctoral Fellowship 1272625N).

\section*{Conflict of interest}

I have no conflict of interest to declare.

\section*{Acknowledgements}

I thank two anonymous reviewers and the associate editor for their constructive
criticism that resulted in substantial improvements of the present article.
I thank Christelle Fraïsse for encouragement and feedback, as well as Quinten
Bafort, Felipe Kauai and Frederik Mortier for comments on an early draft
version.
